---
title: "R Notebook"
output: html_notebook
---


```{r,message=FALSE, warning=FALSE, results='hide'}
rm(list = ls())
# You can use this space to load any packages you want. 
library(tidyverse)
library(dplyr)
library(car)
library (GGally)
library(corrplot)
library(MASS)
library(car)
library(Stat2Data)
library(olsrr)
library(leaps)
library(rgl)
library(leaps)
library(caret)
library(lmtest)
library(nortest)
```

```{r}
getwd()
workdata<-read.csv('data\\working_Barcelona.csv')
```

# Assumption check
```{r fig.height=4}
fullmodel1 <- lm(energy_demand ~ E_1 + E_2 + E_25 + temp + humidity + pressure + wind_speed + rain_duration + factor(day_night) + factor(time_band) + factor(season) + factor(weather_main), data = workdata)
summary(fullmodel1)
res.fullmodel1 <- residuals(fullmodel1) # to obtain residuals
std.res.fullmodel1 <- rstandard(fullmodel1) # to obtain standardized residuals
pred.fullmodel1 <- fitted.values(fullmodel1) # to obtain fitted/predicted values
qqnorm(y = std.res.fullmodel1, main = "Normal Q-Q Plot of Standardized Residuals Fullmodel1",
       xlab = "Theoretical Quantiles", ylab = "Sample Quantiles")
qqline(y = std.res.fullmodel1)
resplotdata1 <- data.frame(std.res.fullmodel1, pred.fullmodel1)
resbf1 <- lm(std.res.fullmodel1 ~ pred.fullmodel1, data = resplotdata1)
plot(x = pred.fullmodel1, y = std.res.fullmodel1, ylab = "Standardized Residuals", xlab = "Predicted Values", main = "Normal Q-Q Plot of Standardized Residuals Fullmodel1", col = ifelse(std.res.fullmodel1 < -3,"red",ifelse(std.res.fullmodel1 > 3,"red","black")))
abline(h = 0, col="blue", lty=1)
abline(resbf1, col="red", lty=3)
abline(h = 3, col="green", lty=3)
abline(h=-3, col="green", lty=3)
legend("bottomleft", legend=c("Best fit line of standardized residuals", "Horizontal line y = 0.0", "Horizontal line, y = +/- 3"), fill = c("red","blue","green"), cex = 1.0)
cat("\n")
vif(fullmodel1)
mclflag1 <- which(vif(fullmodel1) >= 10, arr.ind = TRUE, useNames = TRUE)
ifelse(length(mclflag1) == 0, print(paste("There is no multicollinarity concern with the predictors in this model.")), print(paste("The predictor where there is a multicollinarity concern is", dimnames(mclflag1)[1],".")))
```
Comment(s)/observation(s) - The bulk of the data points in the Standardized residuals vs predicted values plot appear to be clustered between $\sim \pm 3$ but there are are a considerable number (red colored) of potetial outlier points outside this cluster.  Wrt class notes Lesson 6, the items to check on the residuals vs fitted value plot are linearity, non-constant variances and possible outliners.  Here, linearity looks to hold even with the numerous potential outliers to the $\pm3\sigma$ standard deviations.  As heteroscedasticity is not present, I do not see a violation in constant variances.  The bulk of the standardized residuals data appear to be normally distributed based on falling on the Normal Q-Q plot's reference line, though skewness in both tails is still evident.  Given this, I conclude these residuals are aligned with the desired state of "being randomly scattered about the zero line with no apparent pattern." 











