---
title: "501 Project"
output:
  html_notebook: default
  pdf_document: default
---

#  Tree Forest Model to Predict Electricity Demand of Five Cities in Spain
## Links documentation

* [Github Code](https://github.com/ShaileshWasti/STAT-501-Project.git)
* [Kaggle Data](https://www.kaggle.com/datasets/nicholasjhana/energy-consumption-generation-prices-and-weather?select=weather_features.csv)
* [Tree Modeliong](https://towardsdatascience.com/data-science-tutorials-training-a-random-forest-in-r-a883cc1bacd1)

## Install all Packages and Import Library 

```{r}
# packages we may use
#install.packages("r package", repos = "http://cran.us.r-project.org")
#install.packages('plyr', repos = "http://cran.us.r-project.org")
#install.packages("tinytex")
library(tinytex)
#install.packages("dplyr")
library(dplyr)
#install.packages("randomForest")
library(randomForest) 
#urlPackage<-'https://cran.r-project.org/src/contrib/Archive/randomForest/randomForest_4.0-1.tar.gz'
#install.packages(urlPackage, repos=NULL, type="source") 
#install.packages("ranger")
library(ranger)
#install.packages("Metrics")
library(Metrics)
require(devtools)
install_version("randomForest", version = "4.7.1", repos = "http://cran.us.r-project.org")
install.packages("installr")
library(installr)

updateR()
```


## Code for Barcelona

### organize data


```{r}
# import data raw data set
w_BCN_raw <- read.csv("working_Barcelona.csv")
e_BCN_raw <- read.csv("evaluation_Barcelona.csv")
w_BCN <- w_BCN_raw[,-c(1,2)]
e_BCN <- e_BCN_raw[,-c(1,2)]
```

```{r}
# changes of data type
class(w_BCN$day_night) = "category"
class(w_BCN$time_band) = "category"
class(w_BCN$season) = "category"
class(w_BCN$weather_main) = "category"

class(e_BCN$day_night) = "category"
class(e_BCN$time_band) = "category"
class(e_BCN$season) = "category"
class(e_BCN$weather_main) = "category"

# check data types
str(w_BCN)
str(e_BCN) 
summary(w_BCN)

```

### RF modeling

```{r}
# fit random forest model
# default RF model
set.seed(1234)
rf <- randomForest(formula = energy_demand ~ ., data = w_BCN)

# find the num of trees in the forest with smallest MSE
plot(rf)

summary(rf)


```



```{r}

n <- which.min(rf$mse)
n
```

```{r}

testBCN <- rf$mse
testBCN <- as.data.frame(testBCN)

#the slope become flat around 100 in the plot, divide test dataset at 100

testBCN1 <- testBCN$test[c(1:100)]
testBCN2 <- testBCN$test[c(100:493)]

# find the average error value of the two groups

avgBCN1 <- mean(testBCN1)
diffBCN1 <- abs(testBCN1 - avgBCN1) / avgBCN1
mean(diffBCN1)
avgBCN2 <- mean(testBCN2)
diffBCN2 <- abs(testBCN2 - avgBCN2) / avgBCN2
mean(diffBCN2)

#the difference between 100 and 500 is much less than that from 1 to 100

```



```{r fig.height=4}
# modify the model with best mtry and importance of the predictors

mtry <- tuneRF(w_BCN,w_BCN$energy_demand, ntreeTry=100,
              stepFactor=1.5,improve=0.01, trace=TRUE, plot=TRUE)
best.m <- mtry[mtry[, 2] == min(mtry[, 2]), 1]

rf_modified <-randomForest(energy_demand~.,data=w_BCN, mtry = best.m, importance=TRUE,ntree=100)


varImpPlot(randomForest(energy_demand~.,data=w_BCN, mtry = best.m, importance=TRUE,ntree=100), main = "Importance of the varibales", n.var = 5)

```

### check performance

```{r}
# check the time
system.time(randomForest(energy_demand~.,data=w_BCN, mtry = best.m, importance=TRUE,ntree=100))
```


```{r}
# assess the test set performance of rf_modified with Metrics library
predvalue <- predict(rf_modified, e_BCN)
rmse(e_BCN$energy_demand, predvalue)

pdf("BCN1.pdf",         # File name
    width = 7, height = 4, # Width and height in inches
    bg = "white",          # Background color
    colormodel = "cmyk")  # Color model

plot(e_BCN$energy_demand, predvalue, xlab = "Energy Demand of Test Set ", ylab = "Prediction", main = "rF for Barcelona")
grid(10,10)
dev.off()
```

### ranger modeling

```{r}
# create RF model with ranger package
rf_ranger <- ranger(formula = energy_demand ~ ., data = w_BCN)
rf_ranger

```

### check performance

```{r}
# check execution time
system.time(rf_ranger <- ranger(formula = energy_demand ~ ., data = w_BCN))
```


```{r}
# performance
rmse(
  e_BCN$energy_demand, 
  predict(rf_ranger, e_BCN)$predictions
  )

pdf("BCN2.pdf",         # File name
    width = 7, height = 4, # Width and height in inches
    bg = "white",          # Background color
    colormodel = "cmyk")  # Color model

plot(e_BCN$energy_demand, 
  predict(rf_ranger, e_BCN)$predictions, ylab = "Prediction", xlab = "Energy Demand of Test Set", main = "Ranger for Barcelona")
grid(10,10)
dev.off()





```




## Code for Bilbao

### organize data


```{r}
# import data raw data set
w_BLB_raw <- read.csv("working_Bilbao.csv")
e_BLB_raw <- read.csv("evaluation_Bilbao.csv")
w_BLB <- w_BLB_raw[,-c(1,2)]
e_BLB <- e_BLB_raw[,-c(1,2)]
```

```{r}
# changes of data type
class(w_BLB$day_night) = "category"
class(w_BLB$time_band) = "category"
class(w_BLB$season) = "category"
class(w_BLB$weather_main) = "category"

class(e_BLB$day_night) = "category"
class(e_BLB$time_band) = "category"
class(e_BLB$season) = "category"
class(e_BLB$weather_main) = "category"

# check data types
str(w_BLB)
str(e_BLB) 
summary(w_BLB)

```

### RF modeling

```{r}
# fit random forest model
# default RF model
set.seed(1234)
rf2 <- randomForest(formula = energy_demand ~ ., data = w_BLB)

# find the num of trees in the forest with smallest MSE
plot(rf2)

summary(rf2)


```



```{r}

n2 <- which.min(rf2$mse)
n2
```

```{r}

testBLB <- rf2$mse
testBLB <- as.data.frame(testBLB)

#the slope become flat around 100 in the plot, divide test dataset at 100

testBLB1 <- testBLB$test[c(1:100)]
testBLB2 <- testBLB$test[c(100:499)]

# find the average error value of the two groups

avgBLB1 <- mean(testBLB1)
diffBLB1 <- abs(testBLB1 - avgBLB1) / avgBLB1
mean(diffBCN1)
avgBLB2 <- mean(testBLB2)
diffBLB2 <- abs(testBLB2 - avgBLB2) / avgBLB2
mean(diffBLB2)

#the difference between 100 and 500 is much less than that from 1 to 100

```



```{r fig.height=4}
# modify the model with best mtry and importance of the predictors

mtry <- tuneRF(w_BLB,w_BLB$energy_demand, ntreeTry=100,
              stepFactor=1.5,improve=0.01, trace=TRUE, plot=TRUE)
best.m <- mtry[mtry[, 2] == min(mtry[, 2]), 1]

rf2_modified <-randomForest(energy_demand~.,data=w_BLB, mtry = best.m, importance=TRUE,ntree=100)


varImpPlot(randomForest(energy_demand~.,data=w_BLB, mtry = best.m, importance=TRUE,ntree=100), main = "Importance of the varibales", n.var = 5)

```

### check performance

```{r}
# check the time
system.time(randomForest(formula = energy_demand ~ ., 
                   data = w_BLB, ntree = n2, na.action=na.exclude))
```


```{r}
# assess the test set performance of rf_modified with Metrics library
predvalue <- predict(rf2_modified, e_BLB)
rmse(e_BLB$energy_demand, predvalue)

pdf("BLB1.pdf",         # File name
    width = 7, height = 4, # Width and height in inches
    bg = "white",          # Background color
    colormodel = "cmyk")  # Color model

plot(e_BLB$energy_demand, predvalue, xlab = "Energy Demand of Test Set ", ylab = "Prediction", main = "rF for Bilbao")
grid(10,10)
dev.off()



plot(e_BLB$energy_demand, predvalue, xlab = "Energy Demand of Test Set ", ylab = "Prediction", main = "rF for Bilbao")
```

### ranger modeling

```{r}
# create RF model with ranger package
rf2_ranger <- ranger(formula = energy_demand ~ ., data = w_BLB)
rf2_ranger
```

### check performance

```{r}
# check execution time
system.time(rf2_ranger <- ranger(formula = energy_demand ~ ., data = w_BLB))
```


```{r}
# performance
rmse(
  e_BLB$energy_demand, 
  predict(rf2_ranger, e_BLB)$predictions
  )

pdf("BLB2.pdf",         # File name
    width = 7, height = 4, # Width and height in inches
    bg = "white",          # Background color
    colormodel = "cmyk")  # Color model

plot(
  e_BLB$energy_demand, 
  predict(rf2_ranger, e_BLB)$predictions, ylab = "Prediction", xlab = "Energy Demand of Test Set", main = "Ranger for Bilbao"
  )
grid(10,10)
dev.off()


plot(
  e_BLB$energy_demand, 
  predict(rf2_ranger, e_BLB)$predictions, ylab = "Prediction", xlab = "Energy Demand of Test Set", main = "Ranger for Bilbao"
  )
```



## Code for Madrid

### organize data


```{r}
# import data raw data set
w_MDD_raw <- read.csv("working_Madrid.csv")
e_MDD_raw <- read.csv("evaluation_Madrid.csv")
w_MDD <- w_MDD_raw[,-c(1,2)]
e_MDD <- e_MDD_raw[,-c(1,2)]
```

```{r}
# changes of data type
class(w_MDD$day_night) = "category"
class(w_MDD$time_band) = "category"
class(w_MDD$season) = "category"
class(w_MDD$weather_main) = "category"

class(e_MDD$day_night) = "category"
class(e_MDD$time_band) = "category"
class(e_MDD$season) = "category"
class(e_MDD$weather_main) = "category"

# check data types
str(w_MDD)
str(e_MDD) 
summary(w_MDD)

```

### RF modeling

```{r}
# fit random forest model
# default RF model
set.seed(1234)
rf3 <- randomForest(formula = energy_demand ~ ., data = w_MDD)

# find the num of trees in the forest with smallest MSE
plot(rf3)

summary(rf3)


```



```{r}

n3 <- which.min(rf3$mse)
n3
```

```{r}

testMDD <- rf3$mse
testMDD <- as.data.frame(testMDD)

#the slope become flat around 100 in the plot, divide test dataset at 100

testMDD1 <- testMDD$test[c(1:100)]
testMDD2 <- testMDD$test[c(100:462)]

# find the average error value of the two groups

avgMDD1 <- mean(testMDD1)
diffMDD1 <- abs(testMDD1 - avgMDD1) / avgMDD1
mean(diffMDD1)
avgMDD2 <- mean(testMDD2)
diffMDD2 <- abs(testMDD2 - avgMDD2) / avgMDD2
mean(diffMDD2)

#the difference between 100 and 500 is much less than that from 1 to 100

```



```{r fig.height=4}
# modify the model with best mtry and importance of the predictors

mtry <- tuneRF(w_MDD, w_MDD$energy_demand, ntreeTry=100,
              stepFactor=1.5,improve=0.01, trace=TRUE, plot=TRUE)
best.m <- mtry[mtry[, 2] == min(mtry[, 2]), 1]

rf3_modified <-randomForest(energy_demand~.,data=w_MDD, mtry = best.m, importance=TRUE,ntree=100)


varImpPlot(randomForest(energy_demand~.,data=w_MDD, mtry = best.m, importance=TRUE,ntree=100), main = "Importance of the varibales", n.var = 5)

```

### check performance

```{r}
# check the time
system.time(randomForest(formula = energy_demand ~ ., 
                   data = w_MDD, ntree = 100, na.action=na.exclude))
```


```{r}
# assess the test set performance of rf_modified with Metrics library
predvalue <- predict(rf3_modified, e_MDD)
rmse(e_MDD$energy_demand, predvalue)

pdf("MDD1.pdf",         # File name
    width = 7, height = 4, # Width and height in inches
    bg = "white",          # Background color
    colormodel = "cmyk")  # Color model

plot(e_MDD$energy_demand, predvalue, ylab = "Prediction", xlab = "Energy Demand of Test Set", main = "rF for Madrid")

grid(10,10)
dev.off()

plot(e_MDD$energy_demand, predvalue, ylab = "Prediction", xlab = "Energy Demand of Test Set", main = "rF for Madrid")


  

```

### ranger modeling

```{r}
# create RF model with ranger package
rf3_ranger <- ranger(formula = energy_demand ~ ., data = w_MDD)
rf3_ranger
```

### check performance

```{r}
# check execution time
system.time(rf3_ranger <- ranger(formula = energy_demand ~ ., data = w_MDD))
```


```{r}
# performance
rmse(
  e_MDD$energy_demand, 
  predict(rf3_ranger, e_MDD)$predictions
  )

pdf("MDD2.pdf",         # File name
    width = 7, height = 4, # Width and height in inches
    bg = "white",          # Background color
    colormodel = "cmyk")  # Color model

plot(
  e_MDD$energy_demand, 
  predict(rf2_ranger, e_MDD)$predictions, ylab = "Prediction", xlab = "Energy Demand of Test Set", main = "Ranger for Madrid"
  )
grid(10,10)
dev.off()

plot(
  e_MDD$energy_demand, 
  predict(rf3_ranger, e_MDD)$predictions, ylab = "Prediction", xlab = "Energy Demand of Test Set", main = "Ranger for Madrid"
  )
```



## Code for Seville

### organize data


```{r}
# import data raw data set
w_SVL_raw <- read.csv("working_Seville.csv")
e_SVL_raw <- read.csv("evaluation_Seville.csv")
w_SVL <- w_SVL_raw[,-c(1,2)]
e_SVL <- e_SVL_raw[,-c(1,2)]
```

```{r}
# changes of data type
class(w_SVL$day_night) = "category"
class(w_SVL$time_band) = "category"
class(w_SVL$season) = "category"
class(w_SVL$weather_main) = "category"

class(e_SVL$day_night) = "category"
class(e_SVL$time_band) = "category"
class(e_SVL$season) = "category"
class(e_SVL$weather_main) = "category"

# check data types
str(w_SVL)
str(e_SVL) 
summary(w_SVL)

```

### RF modeling

```{r}
# fit random forest model
# default RF model
set.seed(1234)
rf4 <- randomForest(formula = energy_demand ~ ., data = w_SVL)
# find the num of trees in the forest with smallest MSE
plot(rf4)

```

```{r}
n4 <- which.min(rf4$mse)
n4
```


```{r}
# modify the RF model
rf4_modified <- randomForest(formula = energy_demand ~ ., 
                   data = w_SVL, ntree = n4, na.action=na.exclude)
rf4_modified


```

### check performance

```{r}
# check the time
system.time(randomForest(formula = energy_demand ~ ., 
                   data = w_SVL, ntree = n4, na.action=na.exclude))
```


```{r}
# assess the test set performance of rf_modified with Metrics library
predvalue <- predict(rf4_modified, e_SVL)
rmse(e_SVL$energy_demand, predvalue)
plot(e_SVL$energy_demand, predvalue, main = "Seville1")
```

### ranger modeling

```{r}
# create RF model with ranger package
rf4_ranger <- ranger(formula = energy_demand ~ ., data = w_SVL)
rf4_ranger
```

### check performance

```{r}
# check execution time
system.time(rf4_ranger <- ranger(formula = energy_demand ~ ., data = w_SVL))
```


```{r}
# performance
rmse(
  e_SVL$energy_demand, 
  predict(rf4_ranger, e_SVL)$predictions
  )
plot(
  e_SVL$energy_demand, 
  predict(rf4_ranger, e_SVL)$predictions, xlab = "predvalue", main = "Seville2"
  )
```


## Code for Valencia

### organize data


```{r}
# import data raw data set
w_VLC_raw <- read.csv("working_Valencia.csv")
e_VLC_raw <- read.csv("evaluation_Valencia.csv")
w_VLC <- w_VLC_raw[,-c(1,2)]
e_VLC <- e_VLC_raw[,-c(1,2)]
```

```{r}
# changes of data type
class(w_VLC$day_night) = "category"
class(w_VLC$time_band) = "category"
class(w_VLC$season) = "category"
class(w_VLC$weather_main) = "category"

class(e_VLC$day_night) = "category"
class(e_VLC$time_band) = "category"
class(e_VLC$season) = "category"
class(e_VLC$weather_main) = "category"

# check data types
str(w_VLC)
str(e_VLC) 
summary(w_VLC)

```

### RF modeling

```{r}
# fit random forest model
# default RF model
set.seed(1234)
rf5 <- randomForest(formula = energy_demand ~ ., data = w_VLC)
# find the num of trees in the forest with smallest MSE
plot(rf5)

```

```{r}
n5 <- which.min(rf5$mse)
n5
```


```{r}
# modify the RF model
rf5_modified <- randomForest(formula = energy_demand ~ ., 
                   data = w_VLC, ntree = n5, na.action=na.exclude)
rf5_modified


```

### check performance

```{r}
# check the time
system.time(randomForest(formula = energy_demand ~ ., 
                   data = w_VLC, ntree = n5, na.action=na.exclude))
```


```{r}
# assess the test set performance of rf_modified with Metrics library
predvalue <- predict(rf5_modified, e_VLC)
rmse(e_VLC$energy_demand, predvalue)
plot(e_VLC$energy_demand, predvalue, main = "Valencia1")
```

### ranger modeling

```{r}
# create RF model with ranger package
rf5_ranger <- ranger(formula = energy_demand ~ ., data = w_VLC)
rf5_ranger
```

### check performance

```{r}
# check execution time
system.time(rf5_ranger <- ranger(formula = energy_demand ~ ., data = w_VLC))
```


```{r}
# performance
rmse(
  e_VLC$energy_demand, 
  predict(rf5_ranger, e_VLC)$predictions
  )
plot(
  e_VLC$energy_demand, 
  predict(rf5_ranger, e_VLC)$predictions, xlab = "predvalue", main = "Valencia2"
  )
```


